---
title: "exercise-11"
author: "Janny Mauricio Velasco Alban"
format: html
editor: visual
---

# Exercise 11

## Janny Mauricio Velasco Alban

### Practice Model Selection

The objective of this exercise is to run exploratory multivariate regression to evaluate different models.

#### Libraries

```{r}

library(tidyverse)
library(broom)
library(skimr)
library(MuMIn)
library(naniar)
library(ggplot2)
```

#### Dataset

```{r}

f<- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"

d<- read_tsv(f, col_names = TRUE)

view(d)
names(d)

glimpse(d)

s<- (skim(d))
s
```

After doing the skim analysis of the dataset we obtained that 4 columns were categorical (character) and 10 were numeric. (verify with the skimr results).

#### Challenge

##### Step 1

```{r}

# Replace all values of -999 with NA

d<- d %>%
  replace_with_na_all(condition = ~.x == -999)

glimpse(d)
view(d)

# the new dataset only contains NA instead of -999 values.
```

##### Step 2

```{r}

# Drop the variables litter size and refs
# Rename the variable max.life(mo) to maxlife(mo)
# Rename the variable wean mass(g) to weanmass(g)

d<- d %>%
  select(-`litter size`, -refs) %>%
  rename(
    `maxlife(mo)`= `max. life(mo)`,
    `weanmass(g)`= `wean mass(g)`
    
  )

glimpse(d)
view(d)

```

##### Step 3

```{r}

# Log transform all the of the other numeric variables

d<-d %>%
  mutate(across(where(is.numeric), ~ ifelse(.x > 0, log(.x), NA_real_)))

glimpse(d)
view(d)

```

##### Step 4

```{r}

# Regress the age, AFR(mo), and maxlife(mo) and mass variables on overall body mass(g) and add the residuals to the dataframe as new variables

## Age variables regressed on log body mass

m_gest <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)


m_wean <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)


m_afr <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)


m_maxlife <- lm(`maxlife(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)

## Mass variables regressed on log body mass

m_newborn <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)

m_weanmass <- lm(`weanmass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)

## Resigual columns

d <- d%>%
  mutate(
    relGest = resid(m_gest),
    relWean = resid(m_wean),
    relAFR = resid(m_afr),
    relLife = resid(m_maxlife),
    relNewbornMass = resid(m_newborn),
    relWeaningMass = resid(m_weanmass)
  )

glimpse(d)
view(d)

```

##### Step 5

```{r}

 

theme_ada <- theme_bw()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 9, color = "gray40")
  )

# Plot residuals of max lifespan (relLife) in relation to Order

p_life <- ggplot(
  d%>% filter(!is.na(relLife)),
  aes(x = reorder(order, relLife, FUN = median, na.rm = TRUE), y = relLife)
)+
  geom_boxplot(fill = "steelblue", alpha = 0.6, outlier.size = 0.8)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")+
  labs(
    title = "Residual Max Lifespan by Mammalian Order",
    subtitle = "relLife = residuals of log(maxlife) ~ log(mass)",
    x = "Order",
    y = "Residual Max Lifespan"
  )+
  theme_ada

p_life
```

The mammalian order that has the highest residual lifespan is "primates"

```{r}

# Plot residuals of newborn mass (relNewbornMass) in relation to Order

p_newborn <- ggplot(
  d %>% filter(!is.na(relNewbornMass)),
  aes(x = reorder(order, relNewbornMass, FUN = median, na.rm = TRUE), y = relNewbornMass)
)+
  geom_boxplot(fill = "darkorange", alpha = 0.6, outlier.size = 0.8)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")+
  labs(
    title = "Residual Newborn Mass by Mammalian Order",
    subtitle = "relNewbornMass = residuals of log(newborn(g)) ~ log(mass)",
    x = "Order",
    y = "Residual Newborn Mass"
  )+
  theme_ada

p_newborn

```

The mammalian order that has the highest residual newborn mass is Perissodactyla

```{r}

# Plot residuals of wearing mass (relWeaningMass) in relation to Order

p_weanmass <- ggplot(
  d %>% filter(!is.na(relWeaningMass)),
  aes(x = reorder(order, relWeaningMass, FUN = median, na.rm = TRUE), y = relWeaningMass)
)+
  geom_boxplot(fill = "seagreen", alpha = 0.6, outlier.size = 0.8)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")+
  labs(
    title = "Residual Weaning Mass by Mammalian Order",
    subtitle = "relWeaningMass = residuals of log(weanmass(g)) ~ log(mass)",
    x = "Order",
    y = "Residual Weaning Mass"
  )+
  theme_ada

p_weanmass
```

The mammalian order that has the highest residual weaning mass is Pholidota

##### Step 6

```{r}

# Run models and a model selection process to evaluate what variables best predict each of the two response variables, maxlife(mo) and AFR(mo), from the following predictors: gestation(mo), newborn(g), weaning(mo), weanmass(g), litters/year, and overall body mass(g)

##Predictors for model selection

predictor_cols <- c("gestation(mo)", "newborn(g)", "weaning(mo)",
                    "weanmass(g)", "litters/year", "mass(g)")

run_model_selection <- function(response, data) {
  response_clean <- gsub("`", "", response)
  predictor_clean <- c("gestation(mo)", "newborn(g)", "weaning(mo)", "weanmass(g)", "litters/year", "mass(g)")
  
  full_formula <- reformulate(
    termlabels = paste0("`", predictor_clean, "`"),
    response = paste0("`", response_clean, "`"))
  
  vars_bare <- c(
    response_clean, predictor_clean, "`")
  
  data_clean<- data %>%
    drop_na(any_of(vars_bare))
  cat("\nRows used for", response, ":", nrow(data_clean), "\n")
  
  options(na.action = "na.fail")
  full_model<- lm(full_formula, data = data_clean)
  
  dredge_table <- dredge(full_model, rank = "AICc")
  
  return(list(
    full_model = full_model,
    dredge_table = dredge_table,
    data_clean = data_clean
  ))
}

```

```{r}

# Model selection for AFR(mo)

cat("\n========== MODEL SELECTION: AFR(mo) ==========\n")
res_maxlife <- run_model_selection("maxlife(mo)", d)
res_afr <- run_model_selection("AFR(mo)", d)

## the top models

(res_afr$dredge_table)
```

For each of the two response variables, indicate what is the best model overall based on AICc and how many models have a delta AICc?

```{r}

best_afr <- get.models(res_afr$dredge_table, subset = 1)[[1]]
cat("\nBest model for AFR(mo):\n")
print(summary(best_afr))
```

How many models have a delta AICc of 4 or less?

```{r}

top_afr<- get.models(res_afr$dredge_table, subset = delta <= 4)
cat("\nNumber of models with delta AICc <= 4 (AFR):", length(top_afr), "\n")
 
top_afr_table<- res_afr$dredge_table[res_afr$dredge_table$delta <= 4, ]
cat("\nTop models table (AFR):\n")
print(top_afr_table)
 
cols_in_table_afr<- intersect(predictor_cols, colnames(top_afr_table))
in_all_afr <- cols_in_table_afr[
  apply(top_afr_table[, cols_in_table_afr, drop = FALSE], 2,
        function(x) all(!is.na(x)))
]
cat("Variables in ALL top models (AFR):", paste(in_all_afr, collapse = ", "), "\n")

```

What variables, if any, appear in all of this set of "top" models?

Calculate and plot the model-averaged coefficients and their CIs across this set of top models.

```{r}

avg_afr<- model.avg(res_afr$dredge_table, subset = delta <= 4)
cat("\nModel-averaged coefficients (AFR):\n")
print(summary(avg_afr))
 
coef_afr<- as.data.frame(confint(avg_afr)) %>%
  rownames_to_column("term") %>%
  rename(lower = `2.5 %`, upper = `97.5 %`) %>%
  mutate(estimate = coef(avg_afr)[term]) %>%
  filter(term != "(Intercept)")
 
p_avg_afr<- ggplot(coef_afr, aes(x = reorder(term, estimate),
                                   y = estimate))+
  geom_point(size = 3, color = "darkorchid")+
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2,
                color = "darkorchid")+
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")+
  coord_flip()+
  labs(
    title= "Model-Averaged Coefficients: AFR(mo)",
    subtitle= "Top models with delta AICc <= 4",
    x= "Predictor",
    y= "Averaged Coefficient (95% CI)"
  ) +
  theme_ada
 
print(p_avg_afr)
```

#### Results

The best model for maxlife(mo) is the lowest AICc row in dredge table

The best model for AFR(mo) is the lowest AICc row in dredge table
